Accelerated evolution of convective simulations

### 2018-02-23 tex/bvp_paper.tex
third draft

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of natural systems.  For example,
flows in the convection zone of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior.
Explicit timestepping methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

While solutions to the problem of divergent \emph{dynamical} timescales have proven useful,
the difference between convective timescales and thermal timescales, which
characterize system relaxation, remain a significant problem facing studies of convection. 
Solar convection is a prime example of this phenomenon;
dynamical timescales in the solar convective zone are relatively short 
(10 min overturn at solar surface, one month solar
rotation rate) compared to the Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
evolving the thermal structure of the system in a meaningful fashion.
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the ratio of the thermal diffusion timescale to dynamical timescales
grows \cite{anders&brown2017}.
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Increasingly turbulent motions 
require finer grid meshes and smaller timesteps
to capture advective dynamics. Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
while the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at lower Ra as initial conditions for a simulation at a higher
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In certain systems, 
the approximate state of the evolved solution can be estimated. There, an
appropriate set of initial conditions can either be solved for analytically
\cite{couston&all2017} or by using knowledge of Mixing Length Theory or other convective
theories to adjust the initial profile towards the proper adiabatic state \cite{brandenburg&all2005}.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally converged solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}, but it is becoming
increasingly difficult. Such evolution is
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only be reasonably run
for tens to hundreds of buoyancy times \cite{stevens&all2011}, much less the
thousands of buoyancy timescales required for thermal convergence.

In this work, we study a method of achieving accelerated evolution of a
convective simulation through adjusting the thermodynamic state using information 
from resolved, convective dynamics. We couple measurements of the (non-converged)
convective solution with knowledge about energy balances in the eventual solution
to self-consistently adjust the mean thermodynamic profile towards its final state. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare solutions reached through the accelerated evolution
method to those that have been evolved through a full diffusive timescale, and we
examine select simulations at very high Ra which have achieved accelerated evolution. Finally,
in section \ref{sec:extensions}, we discuss extensions of the methods presented here,
and we offer concluding remarks.

@@

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that our fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$, except on the
term where the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, acts in the vertical momentum equation, 
where $\rho = \rho_0(1  - \alpha T_1)$.  
Under these constraints, the equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that the dimensionless unit of
length is the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
by these choices, one time unit is a freefall time ($L_z/v_{ff}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv P / \rho_0 + \phi + |\bm{u}|^2 / 2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1.
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

We study 2D and 3D convection in which the domain is a cartesian box, 
whose dimensionless vertical extent is $z \in [-1/2, 1/2]$, 
and which is horizontally periodic with an extent of $x, y \in [0, \Gamma]$ where $\Gamma = 2$ is the aspect
ratio. 
In the 2D cases, we set $v = \partial_y = 0$.
We specify no-slip, impenetrable boundary conditions at both the top and
bottom boundary and we use mixed thermal boundary conditions, such that
\begin{equation}
u = v = w = 0 \, \, \text{at}\,\,z = \pm 1/2, \qquad T_1 = 0 \,\,\text{at}\,\, z=+1/2, \qquad
\frac{\partial T_1}{\partial z} = 0\,\,\text{at}\,\,z=-1/2.
\label{eqn:bcs}
\end{equation}
For this choice of boundary conditions, the critical value of Ra at which
the onset of convection occurs is Ra$_{\text{crit}} = 1295.78$, and the
supercriticality of a run is defined as $S \equiv \text{Ra}/\text{Ra}_{\text{crit}}$.
Studies of convection which aim to model
astrophysical systems such as stars often employ mixed thermal
boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017},
as we do here; however, our choice of thermal boundary conditions here
reflects the fact that the conditions in Eqn. \ref{eqn:bcs} are the simplest
to implement in the process of accelerated evolution (see section \ref{subsection:ae})
we study here.

@@

\subsection{The method of accelerated evolution}
\label{subsection:ae}
Here we describe a method of Accelerated Evolution (AE). We use this method to quickly evolve the
thermodynamic state of our solutions.  We compare this method to Standard Evolution
(SE), in which we naively evolve the atmosphere for one thermal diffusion time,
$t_\kappa = \mathcal{P}^{-1}$. As Ra increases, SE solutions become intractable, 
while the timeframe of convergence for an AE solution remains nearly constant
(in units of freefall times).
For an example of time saving achieved by using AE, we compare
energy traces at $S = 10^5$ from a SE run in Fig \ref{fig:time_trace}a to an AE run
in Fig. \ref{fig:time_trace}b.

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, where $\angles{}_{x,y}$ represent
a horizontal average, are the basis of the AE method. We measure
both of these quantities early in a simulation, as in Fig. \ref{fig:time_trace}c.
At these early stages in the simulation, these flux profiles are highly asymmetric,
with much more flux leaving through the upper boundary than entering through the
lower boundary as the atmosphere approaches the proper isotherm selected by the
fixed temperature upper boundary condition.
However, by calculating the total flux,
F$_{\text{tot}} = $F$_{\text{conv}} +$ F$_{\text{cond}}$, and then calculating the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
we remove the asymmetry and examine which parts of the atmosphere are dominated
by conduction, and which are not.
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the early thermal boundary layers are roughly the proper
width.  Where $f_{\text{conv}} = 1$, convection dominates all transport, and where
$f_{\text{cond}} = 1$, conduction dominates all transport (in the boundary
layers). Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

@@

In the evolved, time-stationary state, the horizontal- and time-average of
Eqns. (\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}), neglecting terms that
vanish due to symmetry, are
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{conv, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, 
statistically stationary state. Furthermore, under the specification of
F$_{\text{conv, ev}}$ and $\angles{\bm{u}\times\bm{\omega}}_{x,y}$,
\emph{the mean thermodynamic structure of the system} ($\angles{\varpi}_{x,y},\,\angles{T_1}_{x,y}$)
\emph{is fully specified}.

The AE method is thus simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the amount that the
flux in the system needs to be reduced by, as a function of height.
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}$, by $\sqrt{\xi}$, such that the product of all fluctuations
(which carry the convective flux) are diminished by a factor of $\xi$.  We solve
Eqns. (\ref{eqn:bouss_BVP_momentum}-\ref{eqn:bouss_BVP_energy}) with F$_{\text{conv, ev}}$
and $\xi\cdot\angles{\bm{u}\times\bm{\omega}}_{x,y}$ for the mean thermodynamic state,
and then continue evolving in time.  This adjustment of the mean profile and the
diminishing of velocities and fluctuations is the AE method, and it can generally
be applied tens of buoyancy times after the peak of convective transient.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

The AE method specified here quickly converges the fluxes within the
system to within a few percent of the fluxes in a system converged through
SE (Fig. \ref{fig:time_trace}d \& e).

@@

\section{Results}
\label{sec:results}
We study evolved solutions achieved through the SE method from convective onset
up to $S = 10^5$ in 2D and $S = 10^4$ in 3D.  These SE runs are compared to
AE runs spanning from onset up to $S = 10^7$ in 2D and $S = 10^4$ in 3D.
For a full list of simulations, we refer the reader to appendix \ref{appendix:run_table}.

We report the time- and volume-averaged values of select measurements of the
evolved solutions in Fig. \ref{fig:parameter_space_comparison}.  We report
the scaling of heat transport in the evolved solution, as quantified by the
Nusselt number, in Fig. \ref{fig:parameter_space_comparison}a.
The volume averaged Nusselt number is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}}.
\end{equation}
In 3D and in 2D when $S < 10^{3+2/3}$, the evolved system is defined by a clear
value of Nu and the convection reaches a statistically stationary state.
In 2D and at larger values of $S$, the value of Nu oscillates as a function of time
as the plumes in the solution oscillate horizontally.  Our choice of no-slip
boundary conditions prevent the fluid from entering a full-fledged shearing state
\cite{goluskin&all2014}, but the oscillatory motions which do arise cause the
system to vary between periods of low heat transport and high heat transport.
The SE simulations which span up to $S = 10^5$ exhibit the same horizontally
oscillatory motion as the AE solutions for the same initial conditions. The
scaling of the mean value of Nu is roughly $\text{Nu} \propto \text{Ra}^{1/5}$,
weaker than that reported in similar systems with fixed-T and fixed-flux boundary
conditions \cite{johnston&doering2009}.  We attribute this weaker Nu scaling to
the oscillatory nature of the plumes, which may have been avoided by previous studies
using bootstrapping techniques as initial conditions.

In Fig. \ref{fig:parameter_space_comparison}b, we report the rms Reynolds number
in the AE solutions, where
$\text{Re} = \angles{|\bm{u}|} / \mathcal{R}$.  This measure scales roughly as
$\text{Re} \propto \text{Ra}^{0.45}$, and shows little variance with time.

@@

\section{Extensions \& Conclusion}
\label{sec:extensions}
In this work we have used \RB convection as a test case for the AE method, but we argue that
the true power of this method is in its extensions to more complicated studies.
Such extensions are beyond the scope of this paper, but we will briefly describe
areas in which extensions to the AE method should be explored.
We note that a strength of the AE method is its generality.  One
need only derive the steady-state, horizontally-averaged equations for a given
equation set (e.g., Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}))
and couple that information with knowledge of the boundary conditions
and current dynamics (as described in
section \ref{subsection:ae} and appendix \ref{appendix:recipe}).

>>>>

\section{Introduction}
\label{sec:intro}
Natural convection occurs in the presence of disparate timescales which
prohibit numericists from studying realistic models of natural systems.  For example,
flows in the convection zones of stars like the Sun are characteristically low Mach number
(Ma) in the deep interior.
Explicit timestepping methods which are bound by the Courant-Friedrich-Lewy
(CFL) timestep limit must resolve the fastest motions (sound
waves), resulting in timesteps which are prohibitively
small for studies of the deep, low-Ma motions. These systems are numerically
stiff, and the difference between
the sound crossing time and the convective overturn time have made studies of low-Ma stellar
convection difficult. Traditionally, approximations such as
the anelastic approximation, in which sound waves are explicitly filtered out,
have been used to study low-Ma flows \cite{brown&all2010, featherstone&hindman2016}.
More recently, advanced numerical techniques which use implicit or mixed
implicit-explicit timestepping mechanisms have made it feasible to study
convection at low Mach numbers \cite{viallet&all2011, viallet&all2013, viallet&all2016, lecoanet&all2014,
anders&brown2017, bordwell&all2018}, and careful studies of deep convection which
would have been impossible a decade ago are now widely accessible.

Convective systems with divergent dynamical timescales can now be studied; however,
thermal timescales, which characterize system relaxation, are often much larger than
dynamical timescales, and resolving dynamics in atmospheres which are sufficiently
thermally relaxed remains a challenging problem.
Solar convection is a prime example of this phenomenon.
Dynamical timescales in the solar convective zone are relatively short 
(10 min overturn at solar surface, one month solar
rotation rate) compared to the Sun's Kelvin-Helmholtz timescale of
$3 \cdot 10^7$ years \cite{stix2003}.  
In such a system, it is impossible to resolve the convective dynamics while also
evolving the thermal structure of the system in a meaningful fashion using
traditional timestepping techniques alone.
As modern simulations aim to model natural convection
by increasing into the high-Rayleigh-number (Ra) regime,
the thermal diffusion timescale becomes intractably large
compared to dynamical timescales \cite{anders&brown2017}.
Furthermore, as dynamical and thermal timescales separate, 
simulations become more turbulent. Increasingly turbulent motions 
require finer grid meshes and smaller timesteps
to capture advective dynamics. Thus, the progression of simulations into the high-Ra
regime of natural convection is slowed by two simultaneous effects: timestepping
through a single convective overturn time becomes more computationally expensive
and the number of overturn times required for systems to reach thermal equilibration
grows.

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally converged solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  Bootstrapped solutions are succeptible to hysteresis
effects, in which large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
Another commonly-used tactic in
moderate-Ra simulations is to use 
a simple model of the full convective state as initial conditions.  
For example, past studies have used a linear eigenvalue solve to set the initial
convective state \cite{hurlburt&all1984} or used an axisymmetric solution 
as initial conditions for convection in a 3D cylinder \cite{verzicco&camussi1997}. 
In other systems, 
the approximate state of the evolved solution can be estimated. There, a
set of initial conditions which is close to the evolved state be 
derived analytically \cite{couston&all2017, brandenburg&all2005}.

Despite the numerous methods that have been used,
the most straightforward way to achieve a thermally converged solution
is to evolve a convective simulation through a thermal timescale. Some modern
studies do just that \cite{featherstone&hindman2016}.
However, such evolution is
\emph{expensive}, and state-of-the-art simulations at the highest values of Ra
can only reasonably be run
for tens to hundreds of buoyancy times \cite{stevens&all2011}, much less the
thousands of buoyancy timescales required for thermal convergence.

In this work, we study a method of achieving accelerated evolution of
convective simulations. We couple measurements of the dynamics of non-converged
convective simulations with knowledge about energy balances in the desired solution
to self-consistently adjust the mean thermodynamic profile towards its evolved state. 
While such a method has been used previously \cite{hurlburt&all1986}, 
we find no explanation in the current literature of the steps involved in employing
this method, nor any study into the accuracy of such a method.
In section \ref{sec:experiment}, we describe our convective simulations, our
numerical methods, and our method for achieving accelerated evolution. In
section \ref{sec:results}, we compare solutions reached through the accelerated evolution
method to those that have been evolved through a full diffusive timescale, and we
examine select simulations at high Ra which have achieved accelerated evolution. Finally,
in section \ref{sec:extensions}, we discuss extensions of the methods presented here,
and we offer concluding remarks.

@@

\section{Experiment}
\label{sec:experiment}
We study incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that our fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except where it is $\rho = \rho_0(1  - \alpha T_1)$ on the term where
the constant gravitational acceleration, $\bm{g} = - g\hat{z}$, 
acts in the vertical momentum equation.
The equations of motion are \cite{spiegel&veronis1960}
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\hat{x} + v\hat{y} + w\hat{z}$ is the velocity, 
$T = T_0 + T_1$ are the initial and fluctuating components of temperature, 
and $P$ is the pressure.
We non-dimensionalize these equations such that the
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \grad T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \grad T_0}$).
by these choices, one time unit is a freefall time ($L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv P / \rho_0 + \phi + |\bm{u}|^2 / 2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. As $P$ is a
Langrange multiplier under the Oberbeck-Boussinesq approximation, $\varpi$
can be treated straightforwardly as a linear variable. 
In non-dimensional form, Eqns. \ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{gather}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} = \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} = - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{gather}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{P}$ and $\mathcal{R}$ 
are set by the Rayleigh and Prandtl numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \grad T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr$ = 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.

We study 2D and 3D convection in which the domain is a cartesian box, 
whose dimensionless vertical extent is $z \in [-1/2, 1/2]$, 
and which is horizontally periodic with an extent of $x, y \in [0, \Gamma]$,
where $\Gamma = 2$ is the aspect ratio. 
In 2D simulations, we set $v = \partial_y = 0$.
We specify no-slip, impenetrable boundary conditions at both the top and
bottom boundary and we use mixed thermal boundary conditions, such that
\begin{equation}
u = v = w = 0 \, \, \text{at}\,\,z = \pm 1/2, \qquad T_1 = 0 \,\,\text{at}\,\, z=+1/2, \qquad
\frac{\partial T_1}{\partial z} = 0\,\,\text{at}\,\,z=-1/2.
\label{eqn:bcs}
\end{equation}
For this choice of boundary conditions, the critical value of Ra at which
the onset of convection occurs is Ra$_{\text{crit}} = 1295.78$, and the
supercriticality of a run is defined as $S \equiv \text{Ra}/\text{Ra}_{\text{crit}}$.
Studies of convection which aim to model
astrophysical systems such as stars often employ mixed thermal
boundary conditions \cite{hurlburt&all1984, cattaneo&all1991, korre&all2017},
as we do here; however, our choice of thermal boundary conditions here
reflects the fact that the conditions in Eqn. (\ref{eqn:bcs}) are the simplest
to implement in the process of accelerated evolution (see section \ref{subsection:ae})
we study here.

@@

\subsection{The method of Accelerated Evolution}
\label{subsection:ae}
Here we describe a method of Accelerated Evolution (AE), which we use 
to rapidly evolve the thermodynamic state of convective simulations.  
We compare this AE method to Standard Evolution
(SE), in which we naively evolve the atmosphere for one thermal diffusion time,
$t_\kappa = \mathcal{P}^{-1}$. As Ra increases, SE solutions become intractable, 
while the timeframe of convergence for an AE solution remains nearly constant
in freefall time units.
For an example of time saving achieved by using AE, we compare
energy traces at $S = 10^5$ from a SE run in Fig. \ref{fig:time_trace}a to an AE run
in Fig. \ref{fig:time_trace}c.

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\text{cond}} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective flux,
F$_{\text{conv}} = \angles{w(T_0 + T_1)}_{x,y}$, where $\angles{}_{x,y}$ represent
a horizontal average, are the basis of the AE method. We measure
both of these quantities early in a simulation, retrieving profiles similar to
those shown in Fig. \ref{fig:time_trace}b.
At early stages in the simulation, these flux profiles are highly asymmetric,
with more flux exiting the atmosphere at the upper boundary than 
the fixed-flux lower boundary is providing.
By calculating the total flux,
F$_{\text{tot}} =$ F$_{\text{conv}} +$ F$_{\text{cond}}$, we derive the profiles
\begin{equation}
f_{\text{conv}}(z) = \frac{F_{\text{conv}}}{F_{\text{tot}}},\qquad
f_{\text{cond}}(z) = \frac{F_{\text{cond}}}{F_{\text{tot}}},
\label{eqn:bvp_ratios}
\end{equation}
which have the systematic asymmetries removed. These profiles describe which
parts of the atmosphere depend on convection to carry flux (where $f_{\text{conv}}(z) = 1$
and $f_{\text{cond}}(z) = 0$).
We presume that the early convection occupies roughly the same volume as the evolved
convection, and thus that the extent of the early thermal boundary layers 
(where $f_{\text{cond}}(z) = 1$ and $f_{\text{conv}}(z) = 0$) 
will not change significantly over the course of the atmosphere's evolution.
Under this assumption, the proper evolved atmospheric flux profiles
are F$_{\text{conv, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{conv}}$
and F$_{\text{cond, ev}} = \text{F}_{\text{bot}}\cdot f_{\text{cond}}$,
where F$_{\text{bot}} = \mathcal{P}$ is the amount of flux entering the
bottom of the atmosphere.

@@

In a time-stationary state, the horizontal- and time-average of
Eqns. (\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}), neglecting terms which
vanish due to symmetry, are
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{conv, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Convective flows
are perturbations around a thermal profile defined by these equations in the proper evolved, 
statistically stationary state. Furthermore, under the specification of
F$_{\text{conv, ev}}$ and $\angles{\bm{u}\times\bm{\omega}}_{x,y}$,
\emph{the mean thermodynamic structure of the system is fully specified}.

Thus, the AE method is simple: we construct F$_{\text{conv, ev}}$ as described above.
Then we calculate a profile, 
$\xi(z) = \text{F}_{\text{conv, ev}}/\text{F}_{\text{conv}}$, which is the fraction,
as a function of height, by which the flux in the system must be reduced to go from
an initial state (Fig. \ref{fig:time_trace}b) to an evolved state (Fig. \ref{fig:time_trace}d).
We multiply the velocities
and the thermal fluctuations, $T - \angles{T}_{x,y}$, by $\sqrt{\xi}$, such 
that the advected, flux-carrying perturbations in the temperature are appropriately
diminished.  We then solve a boundary value problem including
Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}) using F$_{\text{conv, ev}}$
and $\cdot\angles{\bm{u}\times\bm{\omega}}_{x,y}$, after diminishing it by a factor
of $\xi$. We use mixed thermal boundary conditions, as in Eqn. (\ref{eqn:bcs}),
because $F_{\text{bot}}$ is exactly specified, and the temperature profile is pegged
at a reference value for the AE solve.
For specifics on the precise implementation of the AE method, we refer
the reader to appendix \ref{appendix:recipe}.

The AE method converges atmospheeres from extreme flux disequilibrium states 
(Fig. \ref{fig:time_trace}b) on short dynamical timescales (Fig. \ref{fig:time_trace}c)
to a converged state (Fig. \ref{fig:time_trace}d) whose flux balance is very
close to the balance present in the corresponding SE solution (Fig. \ref{fig:time_trace}e).
While the application of the AE method advances solutions toward the proper solution,
we find that multiple applications of the method is the best means of achieving converged
solutions quickly.

@@

\section{Results}
\label{sec:results}
We study evolved solutions achieved through the standard evolution (SE) method 
from convective onset
up to $S = 10^5$ in 2D and $S = 10^4$ in 3D.  These SE runs are compared to
accelerated evolution (AE) runs spanning from onset up to $S = 10^7$ in 2D and $S = 10^4$ in 3D.
For a full list of simulations, we refer the reader to appendix \ref{appendix:run_table}.

We report the time- and volume-averaged values of select measurements of the
evolved solutions in Fig. \ref{fig:parameter_space_comparison}.
The scaling of heat transport in the evolved solution, as quantified by the
Nusselt number, is shown in Fig. \ref{fig:parameter_space_comparison}a.
The volume averaged Nusselt number is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where $\angles{}$ represent a volume average.
In 2D when $S < 10^{3+2/3}$ and in 3D, the evolved system is defined by a clear
value of Nu and the convective heat transport reaches a temporally stationary state.
In 2D and at larger values of $S$, the value of Nu oscillates as a function of time
due to large horizontal oscillations in the convective plume structures.
Our choice of no-slip
boundary conditions prevent the fluid from entering a shearing state
\cite{goluskin&all2014}, but the oscillatory motions which do arise cause the
system to vary between periods of low heat transport and high heat transport.
The SE simulations which span up to $S = 10^5$ exhibit the same horizontally
oscillatory motion as the AE solutions for the same initial conditions. The
scaling of the mean value of Nu is roughly $\text{Nu} \propto \text{Ra}^{1/5}$,
weaker than that reported in similar systems with fixed-T and fixed-flux boundary
conditions \cite{johnston&doering2009}.  We attribute this weaker Nu scaling to
the oscillatory nature of the plumes, which may have been avoided by previous studies
using bootstrapping techniques as initial conditions.

In Fig. \ref{fig:parameter_space_comparison}b, we report the volume-averaged
RMS Reynolds number in the AE solutions, where
$\text{Re} = \angles{|\bm{u}|} / \mathcal{R}$.  This measure scales roughly as
$\text{Re} \propto \text{Ra}^{0.45}$, and shows little variance with time.

@@

\section{Extensions \& Conclusion}
\label{sec:extensions}
In this work we have used \RB convection as a test case for the AE method, but we argue that
the true power of this technique is in its extensions to more complicated studies.
One need only derive the steady-state, horizontally-averaged equations for a given
equation set (e.g., Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}))
and couple that information with knowledge of the boundary conditions
and current dynamics (as described in
section \ref{subsection:ae} and appendix \ref{appendix:recipe}) to achieve a quickly
converged solution. While in-depth studies of AE extensions are beyond the 
scope of this paper, we will briefly describe
areas in which the AE method should be explored.
